A numerical study of fixation probabilities for strategies in the Iterated Prisoner's Dilemma

### 2017-04-14 tex/main.tex
Move comments from MH to TODOs.

To see if one strategy is not neutral with respect to another, we want to
empirically measure the fixation probability and compare to the neutral rate. To
do this right we need a lot of counts, since we're estimating a binomial
probability p with variance $p(1-p) / k$ and $p$ is close to $1 / N$. To get the
variance small you need something like $k>1000$ observations ( we can work out
the precise requirements).

Note we're not estimating $r$ for each strategy (pair) since we're in a
frequency dependent situation, so we need to look at the population states (1,
N-1) and (N-1, 1) for every pair of strategies, i.e. we can't assume that we're
in a $\rho \leftrightarrow 1-\rho$ symmetry. More precisely, $\rho_{(1, N-1)} !=
1 - \rho_{(N-1, 1)}$ in general. However we can (for fun) compute $r$ from
$\rho$ with Newton's method (it's not easily invertable for $N > 3$), or take a
Bayesian approach on what the distribution of $\rho$ is and then compute a
distribution for $r$ in the usual way.

A nice addition would be, for an interesting combination of strategies, to
measure the fixation value for all $(i, N-i)$ and compare to the above formula
for the value of $r$ derived from the $(1, N-1)$ case. This would show how much
we deviate from frequency independence.
The existing notebook attempts to get at 1 and 2 by looking at the distributions
of fixation probabilities for each strategy -- that's what the box plots for
each N try to visualize for particular N, and the "Player Rankings by Median vs.
Population Size" for how the cooperative strategies become more successful as N
increases. That plot is the main takeaway IMO, and reinforces the "evolution of
cooperation" narrative that's so popular. We can tie back to Press and Dyson
here -- yes, ZD strategies are good Head-to-Head and in small populations, but
they aren't great when the population size gets bigger. How much bigger? Even at
N=4 there is a dramatic decline for ZD-extort. Note that this goes against the
claims of Stewart and Plotkin (they claimed that ZD strategies basically
dominate the Moran process no matter how much memory you allow). This also
matches our tournament results -- ZD strategies win matches but not tournaments.

It would be great to see how the ensemble strategies (meta strategies) fare, if
we don't mind burning the CPU cycles. I left them out of my initial analysis.

More future work: * Mutation -- for mutation we no longer have fixation, rather
a stationary distribution. This may require some more programming to compute
efficiently (perhaps my stationary library). There's a lot of interesting work
to do here.

Ip think we'd want to include a few of the heatmaps in the final section of the
notebook for some interesting cases, like FoolMeOnce, EvolvedLookerUp, etc.
Pushing N higher will make all the plots more interesting. How high we can get
N? I'd really like to get it to N >= 11.

Structure:

\begin{itemize}
    \item General overview of the data obtained;
    \item Inclusion of most of the work in \texttt{Moran.ipynb}.
\end{itemize}

@@

Further Variants (possible additions or future papers): * Noise * Spatial
structure * More than two types in the population * Modified Moran processes
(e.g. Fermi selection with the strength of selection coefficient) * Altered game
matrices

Noise is especially interesting because a lot of the cooperative strategies are
going to appear neutral to each other (since neither will cast a D unprovoked).
A little bit of noise should shuffle the ranks around quite a bit, and show off
the abilities of e.g. OmegaTFT. Might be worth including at least one of the
"Player Rankings by Median vs. Population Size" plots for some value of noise
(such as 0.05).

>>>>

% TODO Compare fixation rates to neutral fixation 1/2 (because we're in the N/2)
% case.

% TODO Discuss effect of size on performance of strategies. Relate to Press and
% Dyson work: "ZD strategies win matches but not tournaments".

% TODO Redo everything with Meta strategies? This was a comment by MH. I suggest
% we leave this for this paper.

% TODO Include discussion about noise.

% TODO Discuss some particular interesting cases; strategies that perform well
% in the overall tournament for v2.8.0 perhaps?

@@

% TODO Write overall conclusion. (Perhaps: weakness of short memory
% strategies?)

Further work:

\begin{itemize}
    \item Spatial structure;
    \item More than two types in the population;
    \item Modified Moran processes (Fermi selection);
    \item Mutation;
\end{itemize}
% TODO Write this section.
